---
title: "MSDS 6306: Doing Data Science Tidy Data Live session Unit 05 assignment"
author: "Will Holt"
date: "October 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Questions
#####Backstory: 
Your client is expecting a baby soon. However, he is not sure what to name the child. Being out of the loop, he hires you to help him figure out popular names. He provides for you raw data in order to help you make a decision.

####1. Data Munging (30 points): Utilize yob2016.txt for this question. This file is a series of popular children's names born in the year 2016 in the United States. It consists of three columns with a first name, a gender, and the amount of children given that name. However, the data is raw and will need cleaning to make it tidy and usable.

#####a. First, import the .txt file into R so you can process it. Keep in mind this is not a CSV file. You might have to open the file to see what you're dealing with. Assign the resulting data frame to an object, df, that consists of three columns with humanreadable column names for each.

```{r}
#Download the file yob2016 locally and make sure the proper directory is established
getwd()
#Make sure the file is within the directory
list.files()
#Load the yob2016.txt file and save as df
df <- read.table(file = "yob2016.txt", sep= ";", header = FALSE)
#rename column names
names(df) <- c("Name", "Sex", "Count")
#Check some observations to ensure loaded correctly
head(df)
```

#####b. Display the summary and structure of df

```{r}
summary(df)
str(df)
#changed the format of NAme from Factor to Character
df$Name <- as.character(df$Name)
```

#####c. Your client tells you that there is a problem with the raw file. One name was entered twice and misspelled. The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y's at the end of the name. Write an R command to figure out which name it is and display it.

```{r}
#The following code should return Fionayyy.
df[grep("yyy$", df$V1), ]
```

#####d. Upon finding the misspelled name, please remove this particular observation, as the client says it's redundant. Save the remaining dataset as an object: y2016

```{r}
y2016 <- df[-212,]
#check to make sure the row has been removed
y2016[210:215,]
#Change Column Names
names(y2016) <- c("Name", "Sex", "Count")
#Change Name column to character
y2016$Name <- as.character(y2016$Name)
#check some of the observations to ensure data looks good
#You can use str and dim for more detail if needed
head(y2016)
```

####2. Data Merging (30 points): Utilize yob2015.txt for this question. This file is similar to yob2016, but contains names, gender, and total children given that name for the year 2015.

#####a. Like 1a, please import the .txt file into R. Look at the file before you do. You might have to change some options to import it properly. Again, please give the dataframe human-readable column names. Assign the dataframe to y2015.

```{r}
#Load the yob2015.txt file and save as y2015
y2015 <- read.table(file = "yob2015.txt", sep= ",", header = FALSE)
#rename column names
names(y2015) <- c("Name", "Sex", "Count")
#changed the format of Name from Factor to Character
y2015$Name <- as.character(y2015$Name)
#Check some observations to ensure loaded correctly
head(y2015)
```

#####b. Display the last ten rows in the dataframe. Describe something you find interesting about these 10 rows.

```{r}
#It's rather odd that the last 10 names all have the same count of 5. One would think this would vary.
tail(y2015, 10)
```

#####c. Merge y2016 and y2015 by your Name column; assign it to final. The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.

```{r}
final <- merge(x = y2016, y = y2015, all = FALSE, by = "Name")
final <- final[complete.cases(final),]
#Check NAs
colSums(is.na.data.frame(final))
```


####3. Data Summary (30 points): Utilize your data frame object final for this part.

#####a. Create a new column called Total in final that adds the amount of children in 2015 and 2016 together. In those two years combined, how many people were given popular names?

```{r}
final$Total <- with(final, Count.x + Count.y)
head(final)
#Sum the Total column to get 11404228
sum(final$Total)
```

#####b. Sort the data by Total. What are the top 10 most popular names?

```{r}
final <- final[with(final, order(-Total)), ]
head(final, 10)
```


#####c. The client is expecting a girl! Omit boys and give the top 10 most popular girl's names.

```{r}
Top10Girls <- final[final$Sex.x == 'F',]
head(Top10Girls, 10)
```

#####d. Write these top 10 girl names and their Totals to a CSV file. Leave out the other columns entirely.

```{r}
FinalTop10Girls <- Top10Girls[1:10,]
row.names(FinalTop10Girls) <- c()
FinalTop10Girls <- within(FinalTop10Girls, rm("Sex.x", "Count.x", "Sex.y", "Count.y"))
write.csv(FinalTop10Girls, "FinalTop10Girls.csv")
#Check to see if file is in directory
list.files()
```


####4. Upload to GitHub (10 points): 
Push at minimum your RMarkdown for this homework assignment and a Codebook to one of your GitHub repositories (you might place this in a Homework repo like last week). The Codebook should contain a short definition of each object you create, and if creating multiple files, which file it is contained in. You are welcome and encouraged to add other files just make sure you have a description and directions that are helpful for the grader.

https://github.com/WillHolt-SMU/Top10GirlNames